\documentclass[preprint]{elsarticle}
%\documentclass{article}

\begin{document}

%
% Autores
%
\title{Learning in Small and Large Ubiquitous Computing Environment}

\author[unisinos]{J.~Barbosa} 
\ead{jbarbosa@unisinos.br}

\author[unisinos]{R.~Hahn} 
\ead{rodrigo.rmh@gmail.com}

\author[unisinos]{A.~Wagner} 
\ead{andre.nho@gmail.com}

\author[unilasalle]{D.N.F.~Barbosa} 
\ead{nice@unilasalle.edu.br}

\author[ufrgs]{C.F.R.~Geyer} 
\ead{geyer@inf.ufrgs.br}

\address[unisinos]{PIPCA/Unisinos, São Leoplodo, RS, Brazil}
\address[unilasalle]{UniLasalle, Canoas, RS, Brazil} 
\address[ufrgs]{UFRGS, Porto Alegre, RS, Brazil}

%
% Abstract
%
\begin{abstract}

GlobalEdu is a generic model created to support learning in ubiquitous
computing environments. This model has the necessary support to implement
learning-related functionalities in ubiquitous environments. The basic
ubiquitous computing support must be supplied by a middleware where GlobalEdu
lays atop. This paper1 proposes the integration of GlobalEdu with two
middlewares: ISAM and LOCAL. ISAM supports the creation of large-scale
ubiquitous systems. As such, its integration with GlobalEdu results in
large-scale ubiquitous learning environments. On the other hand, LOCAL is
dedicated to create small-scale, location and context-aware ubiquitous
learning environments. The integration between GlobalEdu and LOCAL results in
a local ubiquitous learning system. We created a practical scenario, where our
proposal was evaluated.

\end{abstract}

\maketitle


%
% Texto
%
\section{Introduction}

Mobile computing \citet{lamport94} is a computational paradigm emerging from wireless networks
and distributed systems technologies. Mobility allows the existence of various
scenarios, from nomadic computing to ubiquitous computing [1, 2]. In the last
scenario, the users, carrying portable devices (such as handhelds and
notebooks) have access to a shared infrastructure, independently from their
physical location or displacement form. In the viewpoint of ubiquitous
computing, the computational power will be available anytime and everywhere. 

In this scenario, new learning perspectives are essential, due to the
relationship between teaching and learning changes [3]. In our point of view,
a great amount of educational resources will soon be distributed in a global
network, and learning processes will need to explore this diversity
everywhere, anytime and in any device the learners may be employing in their
activities. Educational applications will explore all the learning
opportunities, joining contextual information (for example, a work of art in a
museum that is being visited) with educational objectives (the learner can be
studying history of art). 

GlobalEdu is a model oriented to ubiquitous learning. It supplies the
necessary support to explore the new learning opportunities available in
ubiquitous environments. GlobalEdu is structured into layers (detailed in
section 3) and can also be coupled with an ubiquitous computing middleware
(considered the third layer). This paper proposes the integration of GlobalEdu
with two middleware projects: ISAM [4, 5] and LOCAL [6, 7]. The ISAM project
focuses on the building and management of large-scale ubiquitous computing
environments [8], and is being developed at UFRGS (Federal University of Rio
Grande do Sul, Brazil). On the other hand, LOCAL aims to support small-scale,
location-and-context-aware computing environments. It is being developed at
UNISINOS (University of Vale do Rio dos Sinos, Brazil).

% TODO - reorganizar seções
This paper is organized as follows. Section 2 presents the definitions and
standards associated with ubiquitous learning. Section 3 details GlobalEdu.
Section 4 proposes the integration between GlobalEdu and the middleware
projects ISAM and LOCAL. The fifth section describes a practical approach
where a small-scale scenario is being evaluated. Section 6 compares our
proposal with others in the same area. Finally, section 7 presents some
conclusions and plans for future work.

%
% Bibliografia
%
\begin{thebibliography}{9}

\bibitem{lamport94}
  Leslie Lamport,
  \emph{\LaTeX: A Document Preparation System}.
  Addison Wesley, Massachusetts,
  2nd Edition,
  1994.

\end{thebibliography}

\end{document}
